---
title: "Billboard Top 100 Analysis 1965-2015"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: fill
---
This dataset is called [Billboard 1964 – 2015 Songs and Lyrics](https://www.kaggle.com/rakannimer/billboard-lyrics). The Billboard 100 ranks relevant music of the time based on radio plays, streams, and other measurements that ultimately tell how popular a song is at any given time. This data is comprised of the year’s end data for the last 50 years. The final chart is calculated using an inverse point system based on the weekly Billboard charts (100 points for a week at number one, 1 point for a week at 100). This allows for an overall score to be complied to see what artists have really left their mark on that respective year and over the last 50 years. My purpose here is too
see the artist with largest impact over the last 50 years along with the evolution of the songs along the way.
```{r setup, include=FALSE}
library(flexdashboard)
#Setting up workspace
suppressWarnings(library(stringr))
suppressWarnings(library(stringi))
suppressWarnings(library(ggplot2))
suppressWarnings(library(sqldf))
suppressWarnings(library(dplyr))
suppressWarnings(library(tidyr))
suppressWarnings(library(plotly))

#read in data
dataset = read.csv('billboard_lyrics_1964-2015.csv')
str(dataset)
head(dataset)

#data cleansing and exploration
na_count <- colSums(is.na(dataset))
na_count <- data.frame(na_count)
na_count

#clear NAs as they are in the lyrics column therefore not useful for our analysis
df<-na.omit(dataset)

#To use stringi to measure the length of the songs, we need to make sure our 
#string data is in character format
df$Lyrics<-as.character(df$Lyrics)
df$Artist<-as.character(df$Artist)
df$Song<-as.character(df$Song)

#getting song length for by year aggregation
df$song_length <- stri_length(df$Lyrics)

#We first want to clean the data to have just the 
#main artists in the artists column
df$Artist <- str_split_fixed(df$Artist, "featuring", 1)
#Lets look at what artists show up the most
df$Artist<-as.factor(df$Artist)
top_artists <- data.frame(table(df$Artist))
colnames(top_artists) <- c('Artist', 'Appearances')
top_artists$Appearances<- as.numeric(top_artists$Appearances)
top_artists <- top_artists[order(-top_artists$Appearances),]

#create dataframe with top 20 artists based on Billboard apperances
top_20 <- data.frame(top_artists[1:20, ])
colnames(top_20) <- c('Artist', 'Appearances')
top_20$Artist <- as.character(top_20$Artist)
top_20$Artist <- factor(top_20$Artist, levels=unique(top_20$Artist))
#lets look at the top ten songs over the last 50 years
query <- "select * from df where Rank >=1 and Rank <=10"
top_10 <- sqldf(query)

#lets look at who has been in the top 10 most over the last 50 years
top_10$Artist<-as.factor(top_10$Artist)
top_10_artists <- data.frame(table(top_10$Artist))
colnames(top_10_artists) <- c('Artist', 'Appearances')
top_10_artists$Appearances<- as.numeric(top_10_artists$Appearances)
top_10_artists <- top_10_artists[order(-top_10_artists$Appearances),]
top_10_artists <- top_10_artists[1:10, ]
top_10_artists$Artist <- as.character(top_10_artists$Artist)
top_10_artists$Artist <- factor(top_10_artists$Artist, levels=unique(top_10_artists$Artist))

#using sqldf to aggregate data 
query <- "select year, avg(song_length) as avg_song_length from df group by year"
song_length <- sqldf(query)
```

Row
-----------------------------------------------------------------------

### Let's look at song length over the years.

```{r}
#plot results in time series
library(plotly)
plot <- ggplot()+
  geom_line(data=song_length, aes(x=Year, y=avg_song_length),color='blue')+
  geom_point(data=song_length, aes(x=Year, y=avg_song_length),color='red')+
  ggtitle("Average Song Length 1965-2015 (Count by Indiviual Letters)")+
  xlab("Year")+
  ylab("Char. Count")
plot
p1 <- ggplotly(plot)
p1
```

Row
-----------------------------------------------------------------------

### All I Want For Chirstmas Is Top 10 Songs!

```{r}
#lets look at the top artists with the most top ten appearances over the years
#We can now plot the top 20 artists with the most apperances on the Billboard Top 100
plot <- ggplot(data=top_10_artists, aes(x=Artist, y=Appearances, fill=Appearances))+
  geom_bar(stat='identity')+
  xlab('Apperances On The Top 100')+
  ylab('Artists')+
  ggtitle('Top 10 Artists With The Most Top Ten Billboard Songs')+
  coord_flip()
plot
p4 <- ggplotly(plot)
p4
```

### Madonna has left her mark on the music industry.

```{r}
 #We can now plot the top 20 artists with the most apperances on the Billboard Top 100
plot <- ggplot(data=top_20, aes(x=Artist, y=Appearances, fill=Appearances))+
  geom_bar(stat='identity')+
  xlab('Apperances On The Top 10')+
  ylab('Artists')+
  ggtitle('Top 20 Artists With The Most Billboard Top 100 Songs')+
  coord_flip()
plot
p3 <- ggplotly(plot)
p3
```

